Learning constraints in tabular data

### 2016-05-13 Paper/ECAI/tabular.tex
surprisingly difficult reworking of 4.2.3, satisf subgroups

%Due to the complexity of numeric computations in spreadsheets, people often fail to grasp the properties of the data, which leads to the so called Spreadsheet risk\footnote{\url{https://en.wikipedia.org/wiki/Spreadsheet\#Spreadsheet_risk}} and even caused major flaw in the famous economics papers \cite{flaw_excel} and billion-losses in the financial industry \cite{spreadsheet_risk_loss}. Tabular constraint learning provides a potential cure by indicating learned constraint violations and suggest possible fixes.

@@

\subsubsection{Constraints over satisfying subgroups}
\tias{I have not yet reworked this part}

\tias{can implement in any way, CSP or imperative (we: imperative)}
$\findassignment(t,a,C_{solutions})$ finds, for all assignments~$a = (G_1, ..., G_n)$ in $A$, the subgroups $G_i' \subseteq G_i$ such that the corresponding constraint $t(G_1', ..., G_n')$ is satisfied (i.e. the \CSignature and \CFunction of~$t$ are satisfied).
For every \template~$t$ an implementation (declarative or imperative) needs to be provided that, given an assignment, finds valid subgroups.

\paragraph{Example}
Consider a group assignment $G_x, G_y$ associated with the arguments $v_x,v_y$ of \ecrank{v_y}{v_x}, then \findassignment selects single vectors $V_y \in G_y, V_x \in G_x$ such that \ecrank{V_y}{V_x} is true.
For example, in Figure~\ref{fig:main_example} the group $G = \range{T_1}{\rangeall}{\rangeto{3}{8}}$ can serve as both $G_x$ and $G_y$.
Then \findassignment could select $V_x$ to be $\range{T_1}{\rangeall}{7}$ and $V_y$ to be $\range{T_1}{\rangeall}{8}$ since \ecrank{\range{T_1}{\rangeall}{7}}{\range{T_1}{\rangeall}{8}} is satisfied in the data.
\\\\
For some constraint templates (most notably aggregates) there may be multiple (overlapping) subgroups of a given group~$G$ for which the template satisfied.
\tias{like min/max? should mention this example in implied constraints. Not sure we should mention this here, how do we remove overlapping, in post-processing? if so, move to later...}
In this case we choose to omit those subgroups that are fully contained within a larger subgroup since this can lead to over-fitting.
If the returned subsets are too large and accidentally include unrelated vectors the user can split such a group.

\paragraph{Example 2}
Sum Product

Implementation: finds subgroups based on assignments\\
Method $Test(G_R, G_{O_1}, G_{O_2})$\\
$If \lnot G_{O_1} < G_{O_2}: Reject~\#~Canonicity$\\
$If data(G_R) = data(G_{O_1}) \bullet data(G_{O_2}): Accept$
\\\\
Method $FindSubgroups(Assignments)$\\
$Return \{ (G_R', G_{O_1}', G_{O_2}') | \exists (G_R, G_{O_1}, G_{O_2}) \in Assignments \land G_R' \subseteq G_R \land G_{O_1}' \subseteq G_{O_1} \land G_{O_2}' \subseteq G_{O_2} \land Test(G_R', G_{O_1}', G_{O_2}') \}$

Register implementation with the system for constraint X

@@

@@

\begin{table}
  \centering
  \begin{tabular}{lll}
    & \textbf{Total} & \textbf{Average per spreadsheet} \\
    \textbf{Constraints} & $93$ & $2.91$ \\
    \textbf{Accuracy} & $96.77\%$ & $94.27\%$ \\
    \textbf{Redundant} & $23.14\%~(28)$ & $8.33\%~(0.88)$ \\
    \textbf{Speed (s)} & ${\runtotal}s \pm {\runtotalstd}s$ & ${\runfile}s \pm {\runfilestd}s$
  \end{tabular}
  \caption{Overview evaluation on essential constraints of the collected Spreadsheet dataset}
\end{table}
\samuel{Expand summary}
\sergey{would it make sense to specify the size of a spreadsheet in the number of non-zero cells?}

>>>>

\tias{Check that it is clear enough throughout the paper that: our method does not reason at the level of individual cells, but rather about \textit{vectors} of cells, as is common in spreadsheets.}
%Due to the complexity of numeric computations in spreadsheets, people often fail to grasp the properties of the data, which leads to the so called Spreadsheet risk\footnote{\url{https://en.wikipedia.org/wiki/Spreadsheet\#Spreadsheet_risk}} and even caused major flaw in the famous economics papers \cite{flaw_excel} and billion-losses in the financial industry \cite{spreadsheet_risk_loss}. Tabular constraint learning provides a potential cure by indicating learned constraint violations and suggest possible fixes.

@@

\subsubsection{Constraints over satisfying subgroups}
In the previous step, $\generategroups(\textit{t,{\cal G},C})$ generates tuples of candidate groups $(G_1, ..., G_n)$ for the constraint template $t$ with $n$ arguments. In this step, $\findassignment(t, (G_1, \dots, G_n))$, the goal is to enumerate the subgroups $(G'_1, ..., G'_n), \forall i, G'_i \subseteq G_i$ that satisfy both the signature and the definition of the constraint template $t$. The main difference with the previous step is that the group generation reasons about the \textit{properties} of the groups (type, size, length, ...), while this step reasons about the actual \textit{content} of the vectors in the groups, that is, the data.

\paragraph{Example}
Consider the sum-over-rows constraint template $\ecsumr{\sg_r}{\mathbf{\sg_x}}$. In the group generation, all numeric groups where the length of $\sg_r$ vector is equal to the number of rows in $\mathbf{\sg_x}$ has been enumerated. An example from Table~\ref{table:constraints} is $(\sg_r, \mathbf{\sg_x}) = (T_1[:,3:8],T_1[:,3:8])$. Note that this does not match the signature exactly yet: $\sg_r$ contains more than one vector. We can now generate all vectors in $T_1[:,3:8]$ for $\sg_r$ and all subgroups of $T_1[:,3:8]$ for $\mathbf{\sg_x}$ and test whether $\forall i, \sg_r[0][i] = \sum_j \mathbf{\sg_x}[j][i]$ where $G[j][i]$ indicates the $i$the value of the $j$th vector in $G$. An example solution is $(T_1[:7], T_1[:,3:6])$.
\\\\

One can formulate this as a constraint satisfaction problem, similar to what we did in the previous section, or use a simple generate-and-test approach. In contrast to the previous step, the number of subgroups to generate here are typically manageable: $O(n*m^2)$ with $m$ the number of vectors in the largest group; we only consider continuous ranges in subgroups, so a subgroup of size $m$ has $m*(m+1)/2$ possible subgroups to consider. Furthermore, few constraint solvers support floating point numbers, which are prevalent in spreadsheet.

For this reason, our current implementation uses a simple generate-and-test method that takes the required size/length of the groups into account, and sequentially verifies whether the elements of the vectors satisfy the definition. The pseudo-code for this approach is shown in ...\tias{Add pseudocode!!}

\begin{verbatim}
VERY PSEUDOcode:
something recursive that handles vectors/subgroups differently like this?
check(Groups, X, Test):
  if |Groups| == 0:
    forall i in ?: vector elements...
      if not Test(X):
        return []
    return X
  // else
  all = []
  Group = Groups.pop(0)
  if vector(Group):
    forall Vec in Group:
      all += check(Groups, X + Vec, Test)
  else:
    forall Subgroup in Group:
      all += check(Groups, X + Subgroup, Test)
  return all
\end{verbatim}

Note that for some constraint definitions, there are \textit{symmetries} over the arguments that lead to trivially equivalent solutions, for example $G_1 = +(G_2, G_3) \leftrightarrow G_1 = +(G_3, G_2)$. \tias{How are they stopped in the code? should be part of the pseud-code because not part of $Test(X)$?}

\paragraph{Example 2, not used (should be replaced by pseudo-code above?)}
Sum Product

Implementation: finds subgroups based on assignments\\
Method $Test(G_R, G_{O_1}, G_{O_2})$\\
$If \lnot G_{O_1} < G_{O_2}: Reject~\#~Canonicity$\tias{ill-defined, $G_{O_1}$ is a group and $<$ not defined on it (on vector identifiers rather than content?)}\\
$If data(G_R) = data(G_{O_1}) \bullet data(G_{O_2}): Accept$ \tias{does not show that reasons over content of vectors}
\\\\
Method $FindSubgroups(Assignments)$\\
$Return \{ (G_R', G_{O_1}', G_{O_2}') | \exists (G_R, G_{O_1}, G_{O_2}) \in Assignments \land G_R' \subseteq G_R \land G_{O_1}' \subseteq G_{O_1} \land G_{O_2}' \subseteq G_{O_2} \land Test(G_R', G_{O_1}', G_{O_2}') \}$ \tias{does not enforce size of subgroups (e.g. vectors)?}

@@

 Also, there may be overlapping subgroups that satisfy the definition, such as \tias{There is no example in the example figure? I was thinking of max/min, but it is over the entire vector always...}.
In this case we choose to omit those subgroups that are fully contained within a larger subgroup since this can lead to over-fitting.
If the returned subsets are too large and accidentally include unrelated vectors the user can split such a group.

For every \template~$t$ an implementation (declarative or imperative) needs to be provided that, given an assignment, finds valid subgroups.

@@

\begin{table}
  \centering
  \begin{tabular}{lll}
    & \textbf{Total} & \textbf{Average per spreadsheet} \\
    \textbf{Constraints} & $93$ & $2.91$ \\
    \textbf{Accuracy} & $96.77\%$ & $94.27\%$ \\
    \textbf{Redundant} & $23.14\%~(28)$ & $8.33\%~(0.88)$ \\
    \textbf{Speed (s)} & ${\runtotal}s \pm {\runtotalstd}s$ & ${\runfile}s \pm {\runfilestd}s$
  \end{tabular}
  \caption{Overview evaluation on essential constraints of the collected Spreadsheet dataset}
\end{table}
\samuel{Expand summary}
